---
title: "Building KNN Function"
author: "Nouman Riaz"
date: "October 25, 2016"
output: pdf_document
---

#The objective is to implement KNN at 'Default' dataset from ISLR package.

#### First we will load the dataset from ISLR package.

```{r}
require(ISLR)
data("Default")
summary(Default)

```

#### Now we will replace the labels by 'Yes' as '1' and 'No' as '0'

```{r}
Default$default <- as.factor(gsub("Yes","1",Default$default))
Default$default <- as.factor(gsub("No","0",Default$default))
```

#### To split the dataset as 70% training and 30% testing, we will use 'set.seed' function to keep the reproducibility. 

```{r}
smp_size <- floor(0.70 * nrow(Default))

set.seed(123)
train_ind <- sample(seq_len(nrow(Default)), size = smp_size)

train <- Default[train_ind, ]
test <- Default[-train_ind, ]

```

#### We now need to introduce new variables for new labels as per our k-values.

```{r}
test$nlabel_k7 <- as.factor(c(1,0)) #results for k=7
test$nlabel_k11 <- as.factor(c(1,0)) #results for k=11
test$nlabel_k15 <- as.factor(c(1,0)) #results for k=15

```

#### We will now find new labels using KNN technique. That is, first find nearest k-points for the test datapoints and then assign new label as per voting.

#####a) For k=7

```{r}
for(i in 1:nrow(test))
{
  #First we need to calculate distance for each test data point.
  dist <- sqrt((test$income[i]-train$income)^2+(test$balance[i]-train$balance)^2)
  
  ind <- which(dist %in% sort(dist)[1:7]) #finding index of k-nearest points
  labels <- train$default[ind] #Getting labels of nearest points
  test$nlabel_k7[i] <- labels[which.max(labels)] #Selecting the top from voting
}

#Now we will calculate error

error_k7 <- sum(as.numeric(test$default!=test$nlabel_k7))/nrow(test)
cat(error_k7)
```

#####b) For k=11

```{r}
for(i in 1:nrow(test))
{
  #First we need to calculate distance for each test data point.
  dist <- sqrt((test$income[i]-train$income)^2+(test$balance[i]-train$balance)^2)
  
  ind <- which(dist %in% sort(dist)[1:11]) #finding index of k-nearest points
  labels <- train$default[ind] #Getting labels of nearest points
  test$nlabel_k11[i] <- labels[which.max(labels)] #Selecting the top from voting
}

#Now we will calculate error

error_k11 <- sum(as.numeric(test$default!=test$nlabel_k11))/nrow(test)
cat(error_k11)
```

#####c) For k=15

```{r}
for(i in 1:nrow(test))
{
  #First we need to calculate distance for each test data point.
  dist <- sqrt((test$income[i]-train$income)^2+(test$balance[i]-train$balance)^2)
  
  ind <- which(dist %in% sort(dist)[1:15]) #finding index of k-nearest points
  labels <- train$default[ind] #Getting labels of nearest points
  test$nlabel_k11[i] <- labels[which.max(labels)] #Selecting the top from voting
}

#Now we will calculate error

error_k15 <- sum(as.numeric(test$default!=test$nlabel_k15))/nrow(test)
cat(error_k15)
```

