---
title: "Solving Linear Regression Equations"
author: "Nouman Riaz"
date: "September 29, 2016"
output: pdf_document
---

The objective is to find 'Beta' parameters of Linear Regression equation for both Simple and Multiple Linear Regression techniques.

###Part 01 - Simple Linear Regression

The dataset we are using is advertisment dataset from the ISLR website.

```{r}
ads <- read.csv("Advertising.csv", stringsAsFactors=FALSE)
head(ads, n=5) #Observing few dataset entries
```

Now we got to find parameters of two linear equations:

1. Sales ~ Radio
2. Sales ~ Newspaper

```{r, fig.width=6, fig.height=3.7}
#Visualizing these input variables agains the output
plot(ads$Radio, ads$Sales)
plot(ads$Newspaper, ads$Sales)

```

Now solving the following equations to estimate our Beta Values

$$\beta_1 =\frac{\sum_{i=1}^n (x_i - x^-)(y_i - y^-)}{\sum_{i=1}^n (x_i - x^-)^2}$$

$$\beta_0 = y^- - \beta_1 x^-$$

1. Sales ~ Radio

```{r}

x_bar = sum(ads$Radio)/length(ads$Radio) #Calculating xbar, i.e. mean of x
y_bar = sum(ads$Sales)/length(ads$Sales) #Calculating ybar, i.e. mean of y

t = ads$Radio - x_bar #Solving numerator of the equation
s = ads$Sales - y_bar #Solving denominator of the equation

beta_1 <- sum(t*s)/sum(t*t) #Calculating Beta1
beta_0 <- y_bar - (beta_1 * x_bar) #Calculating Beta0

cat(sprintf("Beta_one = %s\n",beta_1))
cat(sprintf("Beta_zero = %s",beta_0))

```

Now calculating $\beta$ values for second variable

2. Sales ~ Newspaper
```{r}
x_bar = sum(ads$Newspaper)/length(ads$Newspaper) #Calculating xbar, i.e. mean of x
y_bar = sum(ads$Sales)/length(ads$Sales)#Calculating ybar, i.e. mean of y

t = ads$Newspaper - x_bar #Solving numerator of the equation
s = ads$Sales - y_bar #Solving denominator of the equation

beta_1 <- sum(t*s)/sum(t*t) #Calculating Beta1
beta_0 <- y_bar - (beta_1 * x_bar) #Calculating Beta0

cat(sprintf("Beta_one = %s\n",beta_1))
cat(sprintf("Beta_zero = %s",beta_0))


```

###Part 02 - Multiple Linear Regression

i.e. Sales ~ TV+Radio+Newspaper

```{r}
library(MASS)
X <- matrix(c(seq(from=1,to=1,length.out = nrow(ads)),ads$TV,ads$Radio,ads$Newspaper), nrow=nrow(ads), ncol=4) #Constructing input matrix
Xt <- t(X) #Taking transpose of input matrix
y <- ads$Sales #Output variable

a <- Xt %*% X #Multiplaying input matrix with its transpose
b <- Xt %*% y #Multiplaying input matrix with output variable
inv <- ginv(a) #Taking inverse
BetaV <- inv %*% b #Caculating Beta vector
cat(BetaV)



```

Verifying

```{r}
model <- lm(formula=Sales~TV+Radio+Newspaper,data=ads)
summary(model)

```

We can see that the 'Estimate' parameter of our Model is exactly same as our calculated Beta values.

